Rényi Entropy for Heterogeneity Identification in SAR Data

### 2025-03-29 Text/Reports/report-Renyi/Appendix3.Rmd
Update corrected comments

>>>>

For a continuous random variable \( Z \) with probability density function (pdf) \( f(z) \), the Rényi entropy of order \( \lambda \in \mathbb{R}_+ \setminus \{1\} \) is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}
Using \eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the \( \Gamma_{\mathrm{SAR}} \) and \( \mathcal{G}^0_I \) distributions:

\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr].
\end{multline}

Similarly, for the \( \mathcal{G}^0_I \) distribution, we obtain:

\begin{multline}
\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr) \\
+ \ln(-1 - \alpha) \\
+ \frac{1}{1 - \lambda} \Bigl[ 
    \lambda\bigl(\ln\Gamma(L - \alpha) - \ln\Gamma(-\alpha)\bigr) \\
    + \ln\Gamma\bigl(\lambda(L - 1) + 1\bigr) 
\Bigr].
\end{multline}

The detailed derivation of the Rényi entropy for the \( \mathcal{G}^0_I \) distribution is provided in the Appendix.

Figure~\ref{fig:entropy_plot} illustrates the behavior of the Rényi entropy as a function of \( \lambda \) for different values of the distribution parameters.

For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$ is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}

Using \eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and $\mathcal{G}^0_I$ distributions. Importantly, the $\mathcal{G}^0_I$ entropy can be expressed in terms of the $\Gamma_{\mathrm{SAR}}$ entropy plus additional terms that account for the roughness parameter $\alpha$:

\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr],
\end{multline}

\begin{multline}
\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) + \Delta H_\lambda(\alpha) \\
\text{where } \Delta H_\lambda(\alpha) = \ln(-1 - \alpha) + \frac{1}{1 - \lambda}
\Bigl[
   \lambda\bigl(
      \ln\Gamma(L - \alpha)
      -\ln\Gamma(-\alpha)
   \bigr) \\
   +\ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)
   -\ln\Gamma\bigl(\lambda(L - \alpha)\bigr)
   +\bigl(\lambda(L-1)+1\bigr)\ln\lambda
\Bigr].
\end{multline}

The term $\Delta H_\lambda(\alpha)$ quantifies the additional entropy contribution due to texture variations in the $\mathcal{G}^0_I$ model. As $\alpha \to -\infty$, we can show that $\Delta H_\lambda(\alpha) \to 0$, recovering the $\Gamma_{\mathrm{SAR}}$ case. This limiting behavior confirms that the $\Gamma_{\mathrm{SAR}}$ distribution is indeed a special case of the $\mathcal{G}^0_I$ distribution for homogeneous areas.

@fig-plot presents the Rényi entropy of the $\mathcal{G}^0_I$ model as a function of $\mu$ for different values of $\alpha$, demonstrating this convergence.

For a continuous random variable \(Z\) with pdf \(f(z)\), the Rényi entropy of order \(\lambda \in \mathbbm{R}_+ \setminus \{1\}\) is defined as:  
\begin{equation}  
\label{E:entropy2}  
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.  
\end{equation}

Using \eqref{E:entropy2}, we derive the Rényi entropy for the \(\Gamma_{\text{SAR}}\) distribution:  
\begin{multline}  
\label{eq-HGammaSAR}  
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) = \ln\mu - \ln L \\  
+ \frac{1}{1-\lambda} \Bigl[ -\lambda\ln\Gamma(L) + \ln\Gamma\bigl(\lambda(L-1)+1\bigr) - \bigl(\lambda(L-1)+1\bigr)\ln\lambda \Bigr].  
\end{multline}

For the \(\mathcal{G}^0_I\) distribution, the entropy **explicitly extends** \(H_\lambda(\Gamma_{\text{SAR}})\) with terms depending on the roughness parameter \(\alpha\):  
\begin{multline}  
\label{eq-HGI0}  
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) + \ln(-1 - \alpha) \\  
+ \frac{1}{1 - \lambda} \Bigl[ \lambda\bigl(\ln\Gamma(L - \alpha) - \ln\Gamma(-\alpha)\bigr) + \ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr) \\  
- \ln\Gamma\bigl(\lambda(L - \alpha)\bigr) + \bigl(\lambda(L-1)+1\bigr)\ln\lambda \Bigr].  
\end{multline}

As \(\alpha \to -\infty\), the additional terms vanish, reducing \eqref{eq-HGI0} to \eqref{eq-HGammaSAR}. This confirms that \(\Gamma_{\text{SAR}}\) is the limiting case of \(\mathcal{G}^0_I\) for homogeneous areas (fully developed speckle).

These results are proven in the Appendix. Notably, the \(\mathcal{G}^0_I\) entropy converges to the \(\Gamma_{\text{SAR}}\) case as \(\alpha \to -\infty\), i.e.,
\[
\lim_{\alpha \to -\infty} H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr).
\]
This limit is analytically derived in the Appendix and corroborates that \(\Gamma_{\text{SAR}}\) describes homogeneous areas (fully developed speckle), while \(\mathcal{G}^0_I\) generalizes it for textured regions.

\section*{Appendix: Proof of the Limit Behavior}  
The limit \(\lim_{\alpha \to -\infty} H_\lambda(\mathcal{G}^0_I) = H_\lambda(\Gamma_{\text{SAR}})\) follows from analyzing the additional terms in \eqref{eq-HGI0}. Let:  
\[
\Delta(\alpha) = H_\lambda(\mathcal{G}^0_I) - H_\lambda(\Gamma_{\text{SAR}}) = \ln(-1-\alpha) + \frac{1}{1-\lambda} \ln\!\left[ \frac{\Gamma(L-\alpha)^\lambda \Gamma(\lambda(-\alpha+1)-1) \lambda^{\lambda(L-1)+1}}{\Gamma(-\alpha)^\lambda \Gamma(\lambda(L-\alpha))} \right].
\]  
Using asymptotic properties of the Gamma function for large \(|\alpha|\):  
\[
\frac{\Gamma(L-\alpha)}{\Gamma(-\alpha)} \sim |\alpha|^L, \quad \frac{\Gamma(\lambda(-\alpha+1)-1)}{\Gamma(\lambda(L-\alpha))} \sim (\lambda|\alpha|)^{\lambda-1 - \lambda L},
\]  
we obtain \(\Delta(\alpha) \sim \ln|\alpha| - \ln|\alpha| = 0\). Thus, \(\Delta(\alpha) \to 0\) as \(\alpha \to -\infty\).

### 2025-03-29 Text/Reports/report-Renyi/R1-report-renyi.Rmd
Update corrected comments

set.seed(1234567890, kind = "Mersenne-Twister")
# Parameters
sample_sizes <- c(49)
R <- 500# Number of replications con 200 y 
B <- 100      # Number of bootstrap replications 10 tambien funciona 0.9 el valor optimo
mu <- 1        # Fixed mu
L <- 1        # Fixed L

@@

#lambda_values <- c(0.8, 0.84, 0.85, 0.86, 0.88, 0.90, 0.91, 0.92, 0.93, 0.94, 0.95, 0.96,0.97,  0.98, 0.99, 1.1, 1.2, 1.3,  1.4, 1.5, 1.6, 1.7, 1.8, 1.9, 2.0, 2.1, 2.2, 2.3, 2.4, 2.5, 3, 3.5 )
#lambda_values <- c(0.65, 0.66, 0.67, 0.68, 0.69, 0.70, 0.71, 0.72, 0.75, 0.9, 0.95, 0.96,0.97,  0.98, 0.99, 1.1, 1.2, 1.3,  1.4, 1.5, 1.6, 1.7, 1.8, 1.9, 2.0 )
lambda_values <- c(  0.9, 0.95,  0.99, 1.1, 1.2,  1.5,  2, 3, 3.5, 4)

>>>>

set.seed(1234567890, kind = "Mersenne-Twister")
# Parameters
sample_sizes <- c(49)
R <- 2000# Number of replications con 200 y 
B <- 300      # Number of bootstrap replications 10 tambien funciona 0.9 el valor optimo
mu <- 1        # Fixed mu
L <- 18        # Fixed L

@@

#lambda_values <- c(0.8, 0.84, 0.85, 0.86, 0.88, 0.90, 0.91, 0.92, 0.93, 0.94, 0.95, 0.96,0.97,  0.98, 0.99, 1.1, 1.2, 1.3,  1.4, 1.5, 1.6, 1.7, 1.8, 1.9, 2.0, 2.1, 2.2, 2.3, 2.4, 2.5, 3, 3.5 )
#lambda_values <- c(0.65, 0.66, 0.67, 0.68, 0.69, 0.70, 0.71, 0.72, 0.75, 0.9, 0.95, 0.96,0.97,  0.98, 0.99, 1.1, 1.2, 1.3,  1.4, 1.5, 1.6, 1.7, 1.8, 1.9, 2.0 )
lambda_values <- c(  0.8, 0.9, 0.95,  0.99, 1.1, 1.2)
